Appendix B: Examples of Deep RL in Ecological Decision Problems

### 2021-06-07 appendices/appendix-B-examples.Rmd
td3-fishing overfish slightly

```{r}
# add a column indicating the model
algos <- c("Optimal", "SAC")
sims_df <- opt_sims %>% mutate(model = factor("Optimal", algos))
policy_df <- opt_policy  %>% mutate(model = factor("Optimal", algos))

@@

We will compare this optimal solution with results from the several different RL algorithms.
Unlike the optimal solution, the RL methods never know the underlying growth function used by the simulation (or even if such a function actually exists).
These methods merely seek to learn a strategy for setting harvest quotas which maximizes the cumulative reward the receive from the environment.
At the heart of each deep RL algorithm is a neural network or a collection of neural networks. The RL agent uses these networks to approximate a policy function and/or a value function. The policy function is the agent's mapping of states to actions -- i.e. the agent's strategy. Value functions attempt to evaluate the goodness of a state or state-action pair towards achieving the RL objective of maximizing cumulative rewards. 
Confusingly, the parameters of a RL algorithm that detail the overall learning process are called _hyper-parameters_ while the parameters of a RL algorithm that are learned during training are called parameters. For example, the number of layers in a neural network is a hyper-parameter but a weight in the neural net is referred to as a parameter.
Before training an agent, we must first specify the hyper-parameters; the algorithm's parameters are often randomly initialized.
Each algorithm may have different hyper-parameters. `stable-baselines3` provides default values for all hyper-parameters based on the original papers that introduced the corresponding algorithms.
For example, here we will try the Soft Actor-Critic algorithm, SAC, training with the default hyperparameters originally posed in @SAC.  We use SAC to train an agent that has a multi-layer perceptron ('MlpPolicy') neural network, with the default architecture of two 64-neuron layers.

```{r}
env = gym$make("fishing-v1", sigma=0.1)
agent = sb3$SAC('MlpPolicy', env, seed = seed) 
```

@@

```{r}
if(!file.exists("../python/cache/sac.zip")) {

@@

  # Save our trained agent for future use
  agent$save("../python/cache/sac")

@@

Note that while default hyper-parameters provide a useful starting place (particularly when the environment has been suitably normalized, a best-practice we discuss above), better performance can almost always be achieved by _tuning_ the hyper-parameter selection.
This is discussed further below.
Having saved our trained agent to disk (`cache/sac.zip`), we can then re-load this agent for evaluation or to continue training.
Note that a copy of the trained agents are included in the corresponding GitHub repository.

```{r}
# load the trained agent from cache
agent = sb3$SAC$load("../python/cache/sac")
```

@@

```{r}
# stack result simulations with those above and plot:
sims_df <- agent_sims %>% 
  mutate(model = factor("SAC", algos)) %>% 
  bind_rows(sims_df)
policy_df <- agent_policy  %>% 
  mutate(model = factor("SAC", algos))  %>% 
  bind_rows(policy_df)

@@

After an initial over-harvest (top panel), SAC achieves nearly optimal performance as measured by reward over time, (bottom right), or compared to the steady-state management solution.
The policy function (bottom left) explains this behavior.
The RL agent has a nearly optimal policy for stock sizes between zero and just above the $B_{MSY}$ target.
At higher stocks, the agent has a tendency to over-harvest.

In this simple example, SAC is able to achieve nearly-optimal performance (as defined by the reward conditions) without tuning, simply by using hyperparameters originally chosen for different RL tasks.
As noted in appendix C, careful design of the environment, such as rescaling of state and action space, can also improve the ability of a Deep RL algorithm to find an acceptable policy without altering the substance of the challenge.
Even so, our agent still had extensive time to train on this specific environment, a total of 300,000 timesteps or 3000 training episodes.
Sufficient tuning of the hyperparameters to this environment may be able result in better performance than we have seen here.
For more difficult problems, it may be impossible to find an acceptable policy without first tuning candidate approaches to select the best hyperparameters for the environment.

@@

```{r}
# Simulate management under the trained agent
# parameters based on statistical best model estimate, see Memarzadeh & Boettiger 2019
# or Appendix D
agent = sb3$SAC$load("../python/cache/sac")

@@

For simplicity of illustration, we will train only a TD3-based agent on this scenario.

>>>>

```{r}
# add a column indicating the model
algos <- c("Optimal", "TD3", "TD3-tuned")
sims_df <- opt_sims %>% mutate(model = factor("Optimal", algos))
policy_df <- opt_policy  %>% mutate(model = factor("Optimal", algos))

@@

We will compare this optimal solution with results from the several different RL algorithms.
Unlike the optimal solution, the RL methods never know the underlying growth function used by the simulation (or even if such a function actually exists).
These methods merely seek to learn a strategy for setting harvest quotas which maximizes the cumulative reward the receive from the environment.
At the heart of each deep RL algorithm is a neural network or a collection of neural networks. The RL agent uses these networks to approximate a policy function and/or a value function. The policy function is the agent's mapping of states to actions -- i.e. the agent's strategy. Value functions attempt to evaluate the goodness of a state or state-action pair towards achieving the RL objective of maximizing cumulative rewards. 
Confusingly, the parameters of a RL algorithm that detail the overall learning process are called _hyper-parameters_ while the parameters of a RL algorithm that are learned during training are called parameters. For example, the number of layers in a neural network is a hyper-parameter but a weight in the neural net is referred to as a parameter.
Before training an agent, we must first specify the hyper-parameters; the algorithm's parameters are often randomly initialized.
Each algorithm may have different hyper-parameters. `stable-baselines3` provides default values for all hyper-parameters based on the original papers that introduced the corresponding algorithms.
<!-- For example, here we will try the Actor-Critic algorithm A2C, training with the default hyperparameters originally posed in @A2C.-->
We use this algorithm to train an agent that has a multi-layer perceptron ('MlpPolicy') neural network, with the default architecture of two 64-neuron layers.

```{r}
env = gym$make("fishing-v1", sigma=0.1)
agent = sb3$TD3('MlpPolicy', env, seed = seed) 
```

@@

```{r}
if(!file.exists("../python/cache/td3_untuned.zip")) {

@@

  # Save our trained agent for future use
  agent$save("../python/cache/td3_untuned")

@@

Note that while default hyper-parameters provide a useful starting place (particularly when the environment has been suitably normalized, a best-practice we discuss above), better performance can almost always be achieved by _tuning_ the hyper-parameter selection.
This is discussed further below.
Having saved our trained agent to disk, we can then re-load this agent for evaluation or to continue training.
Note that a copy of the trained agents are included in the corresponding GitHub repository.

```{r}
# load the trained agent from cache
agent = sb3$TD3$load("../python/cache/td3_untuned")
```

@@

```{r}
# stack result simulations with those above and plot:
sims_df <- agent_sims %>% 
  mutate(model = factor("RL", algos)) %>% 
  bind_rows(sims_df)
policy_df <- agent_policy  %>% 
  mutate(model = factor("RL", algos))  %>% 
  bind_rows(policy_df)

@@

## After Tuning

```{r}
hyper$action_noise = sb3$common$noise$NormalActionNoise(
        mean=np$zeros(1L),
        sigma= 0.6656948079225263* np$ones(1L))
td3 = sb3$TD3('MlpPolicy',
              env,  
              seed = 42L,
              gamma = 0.995,
              learning_rate = 0.0001355522450968401,
              batch_size = 128L,
              buffer_size = 10000L,
              train_freq = 128L,
              gradient_steps = 128L,
              policy_kwargs = list(net_arch=c(400L, 300L)))
```

```{r}
if(!file.exists("../python/cache/td3_tuned.zip")){
  td3$learn(total_timesteps=300000L)
  td3$save("../python/cache/td3_tuned.zip")
}
```

```{r}
td3_sims = env$simulate(td3, reps = 100L)
td3_policy = env$policyfn(td3, reps = 5L)

# stack result simulations with those above and plot:
sims_df <- td3_sims %>% 
  mutate(model = "TD3_tuned") %>% 
  bind_rows(sims_df)
policy_df <- td3_policy  %>% 
  mutate(model = "TD3_tuned")  %>% 
  bind_rows(policy_df)

In this simple example, the RL agent is able to achieve nearly-optimal performance (as defined by the reward conditions) without tuning, simply by using hyperparameters originally chosen for different RL tasks.
As noted in appendix C, careful design of the environment, such as rescaling of state and action space, can also improve the ability of a Deep RL algorithm to find an acceptable policy without altering the substance of the challenge.
Even so, our agent still had extensive time to train on this specific environment, a total of 300,000 timesteps or 3000 training episodes.
Sufficient tuning of the hyperparameters to this environment may be able result in better performance than we have seen here.
For more difficult problems, it may be impossible to find an acceptable policy without first tuning candidate approaches to select the best hyperparameters for the environment.

@@

@@
